---
title: "Data Reformating"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}
# Packages for the data manipulation
library(dplyr)
library(tidyr)

# to call stats filter -stats::filter()
```


# Reformat Catch Data 

* Remove the "all" column
* Create the species column
    - move from wide to long
* General QA approach

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). urn:node:GOA. df35b.304.2. ](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors=FALSE)

head(catch_original)
```


Remove the `all` and `notesRegCode` columns using `select`

Cmd +shift + m : pipe operator shortcut %>%

```{r}
# or select(-All, -notesRegCode)
catch_long <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% # select only these
  gather(key = "species",value = "catch",Chinook, Sockeye, Coho, Pink, Chum) # or you can do gather(key = "species",value = "catch", -Region, -Year) implying gather except for Region and Year


head(catch_long)
```



```{r}
# Obtain the wide version back
catch_wide <- catch_long %>%
  spread(key = species, value = catch)

head(catch_wide)
```

## Clean up our data

* Rename catch to catch_thousands
* Change catch column to numeric
* Create a new catch column in units number of fish

```{r}
catch_clean <- catch_long %>% 
  rename(catch_thousands = catch) %>% # change the word catch to catch_thousands. next replace the i in catch to 1. see test below
  mutate(catch_thousands = ifelse(catch_thousands == "I",1,catch_thousands)) %>% 
  mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands) # drop catch_thousands

head(catch_clean)
```

<!-- Testing chunk -->

```{r, eval=FALSE,echo=FALSE}
# testing chunk

test <- as.numeric(catch_long$catch)
i <- which(is.na(test) == TRUE) # no need to include the == TRUE
catch_long[i,]

```

## Split -Apply - Combine

* Calculate mean catch by species 

```{r}
species_mean <- catch_clean %>% 
  group_by(species, Region) %>% 
  summarise(catch_mean = mean(catch), numb_obs = n())

head(species_mean)
```
```{r}
# Group by year. Also summary gives you only columns you ask for but mutate gives you more columns
year_mean_sd <- catch_clean %>% 
  group_by(Year) %>% 
  summarise(catch_mean = mean(catch), numb_obs = n(), catch_sd = sd(catch)) %>% 
  arrange(-catch_mean) %>%  # from largest to lowest
  filter(Year >= 1990)

head(year_mean_sd)
```



# Join the Region Definitions



#  Misc. Functions

