---
title: "Geospatial Analysis"
author: "Komi Messan"
date: "February 15, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

install.packages("sf")

```{r}
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(leaflet)
library(scales)
```


```{r}
ak_shp_sf <- read_sf("shapefiles/ak_regions_simp.shp")

head(ak_shp_sf)
```


```{r}
plot(ak_shp_sf)
```

The code below provide the datum and projection

```{r}
st_crs(ak_shp_sf)
```


Now transform the data into another projection. The EPSG code can be found on this website [Here](http://spatialreference.org/ref/epsg/?page=1)

```{r}
ak_shp_sf <- ak_shp_sf %>% 
  st_transform(crs = 3338) # 3338 is EPSG code for a location in Alaska

head(ak_shp_sf)

plot(ak_shp_sf)
```


## Spatial Joins

```{r}
pop <- read.csv("shapefiles/alaska_population.csv")

head(pop)
```

Now turn the above csv file into a shape file. Note that any data file that contains long and lat data can be transform into a shape file for a proper geospatial analysis 


```{r}
pop_sf <- st_as_sf(pop,
                   coords = c("lng","lat"),
                   crs = 4326,
                   remove = F) %>% # crs = coordinate reference system
  st_transform(crs = 3338)

 head(pop_sf)
```


 Now we will do the join of the left and right side (i.e. the population and location)
 
```{r}
pop_joined_sf <- st_join(pop_sf, ak_shp_sf, join = st_within)

plot(pop_joined_sf["region"])
```


## Group by an summarize 

```{r}
pop_region <- pop_joined_sf %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarize(total_pop = sum(population))

head(pop_region)
```


```{r}
ak_pop_sf <- left_join(ak_shp_sf, pop_region)

head(ak_pop_sf)

plot(ak_pop_sf["total_pop"])
```


Now group by the management area and plot it

```{r}
ak_mgmt <- ak_pop_sf %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop), do_union = F)

plot(ak_mgmt["total_pop"])
```


Now this code below show how to write an sf file

```{r}
write_sf(ak_shp_sf, "shapefiles/ak_regions_population.shp")
```


# Static plots

```{r}

ggplot(ak_pop_sf) +
  geom_sf(aes(fill = total_pop)) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma)

```

 Now plot the location of each community

 
 

```{r}
ggplot() +
  geom_sf(data = ak_pop_sf, aes(fill = total_pop)) +
  geom_sf(data = pop_sf, size = 0.5) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma)

```



# Interactive maps


* Create a leaflet projection object
 
```{r}
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```



```{r}
ak_crs <- st_transform(ak_pop_sf,crs = 4326)
pop_crs <- st_transform(pop_sf, crs = 4326)
```


```{r}

pal <- colorNumeric(palette = "Reds", domain = ak_crs$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>% 
  addPolygons(data = ak_crs, 
              fillColor = ~pal(total_pop),
              weight = 1,
              color = "black",
              label = ~region,
              fillOpacity = 1) %>% 
  addCircleMarkers(data = pop_crs,
                   lat = ~lat,
                   lng = ~lng,
                   radius = 2,
                   fillColor = "gray",
                   fillOpacity = 1,
                   weight = 0.25,
                   color = "black",
                   label = ~city)
m

```




