---
title: "Geospatial Analysis"
author: "Komi Messan"
date: "February 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

install.packages("sf")

```{r}
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(leaflet)
```


```{r}
ak_shp_sf <- read_sf("shapefiles/ak_regions_simp.shp")

head(ak_shp_sf)
```


```{r}
plot(ak_shp_sf)
```

The code below provide the datum and projection

```{r}
st_crs(ak_shp_sf)
```


Now transform the data into another projection. The EPSG code can be found on this website [Here](http://spatialreference.org/ref/epsg/?page=1)

```{r}
ak_shp_sf <- ak_shp_sf %>% 
  st_transform(crs = 3338) # 3338 is EPSG code for a location in Alaska

head(ak_shp_sf)

plot(ak_shp_sf)
```


## Spatial Joins

```{r}
pop <- read.csv("shapefiles/alaska_population.csv")

head(pop)
```

Now turn the above csv file into a shape file. Note that any data file that contains long and lat data can be transform into a shape file for a proper geospatial analysis 


```{r}
pop_sf <- st_as_sf(pop,
                   coords = c("lng","lat"),
                   crs = 4326,
                   remove = F) %>% # crs = coordinate reference system
  st_transform(crs = 3338)

 head(pop_sf)
```


 Now we will do the join of the left and right side (i.e. the population and location)
 
```{r}
pop_joined_sf <- st_join(pop_sf, ak_shp_sf, join = st_within)

plot(pop_joined_sf["region"])
```


## Group by an summarize 

```{r}
pop_region <- pop_joined_sf %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarize(total_pop = sum(population))

head(pop_region)
```


```{r}
ak_pop_sf <- left_join(ak_shp_sf, pop_region)

head(ak_pop_sf)

plot(ak_pop_sf["total_pop"])
```


Now group by the management area and plot it

```{r}
ak_mgmt <- ak_pop_sf %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop), do_union = F)

plot(ak_mgmt["total_pop"])
```


Now this code below show how to write an sf file

```{r}
write_sf(ak_shp_sf, "shapefiles/ak_regions_population.shp")
```




