---
title: "Data Visualization"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Reformat the data for plotting
<!-- This create a link to a different link from your page -->

For dydplr/tidyr reference see my [data tidying page](https://komimessan.github.io/arctic-training/data-tyding.html)

[Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. urn:node:KNB. doi:10.5063/F1S46Q6M.](https://knb.ecoinformatics.org/view/doi:10.5063/F1S46Q6M)

```{r}

esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e",method = "libcurl"), stringsAsFactors=FALSE)

```


```{r, warning=FALSE, message=FALSE}
# Packages for the data manipulation
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(leaflet)
library(stringr)
library(DT)

```



* calculate annual escapment by species and region
    + `separate`
    + `group_by`
    + `summarize` using the `sum` function
* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)
    + `filter`

<br/>
<br/> <!-- will create space -->

```{r}
# separate
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year","month","day"),
           sep = "-", remove = F) %>% 
  mutate(year=as.numeric(year)) %>% # Coerce year to numeric type
  group_by(year,Species, SASAP.Region) %>% #group by year, specie and region
  summarise(Count = sum(DailyCount), numb_obs = n()) %>% # sum daily count
  filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum")) 


head(annual_esc)
```




# Generate some static plots

<br/> <!-- will create space -->

```{r}
 ggplot(annual_esc, aes(x=Species, y=Count, fill = SASAP.Region)) +
   scale_fill_discrete() +
  geom_col()

# ggplot(annual_esc) +
#   geom_col(mapping = aes(x=Species, y=Count), fill="blue")
```



```{r}
# setting up a theme 
my_theme <- theme_test() +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))
```


```{r}
ggplot(filter(annual_esc,SASAP.Region == "Kodiak"), aes(x=year,y=Count,color=Species)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = comma) +
  ylab("Escapement (number of fish)") +
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  theme_test() +
  my_theme
```


```{r, fig.width= 8, fig.height=7}
p <- ggplot(annual_esc, aes(x=year, y = Count, color = Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels = comma) +
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 3)+ # default is 5 by 5
  my_theme

p

ggsave("figures/regional_escapment.png",
       plot = p,
       width = 8,
       height = 7,
       units = "in")
```


<br/> <!-- will create space -->

# Generate an interactive map

<br/> <!-- will create space -->

Map the sampling locations for escapement

* Find distinct lat/lon/location combinations

<br/> <!-- will create space -->


```{r}
locations <- esc %>% 
  distinct(Location,Latitude,Longitude) %>% 
  drop_na()

head(locations)

```


```{r, echo=FALSE}

datatable(locations)

```



```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~Location)
```


